---
title: "Lab 02 - Plastic waste"
author: "Fiona Wang"
date: "Jan-28-2025"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

```{r initial ggplot, message=FALSE, warning=FALSE}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) + 
  geom_histogram(binwidth = 0.2)
```

There is one country that stands out, let's filter to find this country!

```{r find outlier}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```
It is Trinidad and Tobago from North America. It is a country that I am not familiar with. It seems like they have high consumption of single-use plastics and limited recycling infrastructure.  

### Exercise 1

Remove this text, and add your answer for Exercise 1 here.

```{r plastic-waste-continent}
# insert code here
```

### Exercise 2

```{r plastic-waste-density}
# insert code here
```

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-mismanaged}
# insert code here
```

```{r plastic-waste-mismanaged-continent}
# insert code here
```

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r recreate-viz}
# insert code here
```


