---
title: "Lab 02 - Plastic waste"
author: "Fiona Wang"
date: "Jan-28-2025"
output: 
  html_document: 
    highlight: tango
    theme: yeti
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

```{r initial ggplot, message=FALSE, warning=FALSE}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) + 
  geom_histogram(binwidth = 0.2)
```

There is one country that stands out, let's filter to find this country!

```{r find outlier}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```

It is Trinidad and Tobago from North America. I didn't expect this result, mainly because it is a country that I am not familiar with. It seems like they have high consumption of single-use plastics and limited recycling infrastructure.  

### Exercise 1 Histogram

There weren't a lot of countries reported on Oceania and South America, so they seem to have less plastic waste. North America has the outlier (Trinidad and Tobago). Most countries' plastic waste per capita is between 0 to 0.5.The tallest bars among these graphs are 0.0-0.2 and 0.2-0.4.

```{r plastic-waste-continent, message=FALSE, warning=FALSE}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap, color = continent)) + 
  geom_histogram(binwidth = 0.2) + 
  facet_wrap(~continent, ncol = 3)
```

Another way of visualizing is using density plots. 
```{r density plots, warning=FALSE}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) + 
  geom_density()
```

Compare distributions across continent by coloring density curves by continent.
```{r compare density, warning=FALSE}
ggplot(data = plastic_waste, mapping = aes(x=plastic_waste_per_cap, 
       color = continent)) +
  geom_density()
```

Fill the curves of the density graph and adjust the transparency.
```{r fill density, warning=FALSE}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, color = continent,
                     fill = continent)) + 
  geom_density(alpha = 0.7)
```

### Exercise 2 Recreate density plot
To compare between mapping and setting/geom, I am trying out the following two ways. First, let's use geom_density to define color and fill, and we use mapping to define alpha. The second way is to put color and fill in mapping while defining alpha in the setting/geom.
```{r plastic-waste-density lower alpha, warning=FALSE}
#The first way
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, alpha = 0.1)) + 
  geom_density(color = "pink", fill = "blue")
#The second way
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, fill = continent)) +
  geom_density(alpha = 0.1)
```

2.2 answer: As we can see, if we define color and fill in geom_density, we cannot distinguish between each continent, as this will be a setting for all of the continents. If we put color and fill in mapping, we can ask R to map each continent using different colors. For alpha, when I put it in the mapping, no transparency took place. Also, we want the same transparency to be applied to each of the continent. This is why we just define alpha in the setting, so it can be applied globally.  

Another way to visualize is using box plots.
```{r box plot, warning=FALSE}
ggplot(data = plastic_waste,
       mapping = aes(x = continent, y = plastic_waste_per_cap)) +
  geom_boxplot()
```

### Exercise 3 Violin 

Both boxplot and violin plot give us outliers. The boxplot gives us the median and quartiles that a violin plot doesn't. A violin plot can give you the distribution of each continent which the boxplot doesn't. 

```{r plastic-waste-violin, warning=FALSE}
ggplot(data = plastic_waste,
       mapping = aes(x = continent, y = plastic_waste_per_cap)) +
  geom_violin()
```

### Exercise 4 Scatterplot

4.1. In general, it seems like the more plastic waste per capita, the more mismanaged plastic waste per capita. If there is a linear relationship, it is a positive one. 

```{r plastic-waste-mismanaged, warning=FALSE}
ggplot(data = plastic_waste,
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```


4.2. As we see from the plot below, all continent has a positive relationship between plastic waste per capita and mismanaged plastic waste per capita. Some have steeper slopes: Asia and Africa, whereas some have smaller slopes:North America, Europe).  

```{r plastic-waste-mismanaged-continent, warning=FALSE}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap,
                     y = mismanaged_plastic_waste_per_cap,
                     color = continent)) +
  geom_point(alpha = 0.5)
```

4.3. When asked to visualize this relationship, I was originally thinking that population might affect plastic waste per capita. That's why plastic waste per capita is the x variable, and the population variables are on the y-axis. Both plots don't appear to be perfect linear relationships. Comparing these two plots, I would say that the coastal population plot appears to be more strongly linearly associated.

```{r plastic-waste-population-total, warning=FALSE}
ggplot(data = plastic_waste, mapping = aes(x = total_pop,
                                           y = plastic_waste_per_cap)) + 
  geom_point()
```

```{r plastic-waste-population-coastal, warning=FALSE}
ggplot(data = plastic_waste, mapping = aes(x = coastal_pop,
                                           y = plastic_waste_per_cap)) +
  geom_point()
```

### Exercise 5

Looking at plot 1 below, the general curve shows that the higher the proportion (from 0 to 1), the more plastic waste per capita. There is not much to comment on specific continent, as they are scattered and hard to keep track of. What's interesting is that some countries have proportions that are higher than 1 which is weird.  

Looking at plot 2 below, this graph shows the trend of each continent. All shows a general increase. However, there also seems to be a decrease in trend for some continents. 

```{r recreate-viz, warning=FALSE}
plastic_waste$coastpror = plastic_waste$coastal_pop/plastic_waste$total_pop

ggplot(data = plastic_waste %>%
         filter(plastic_waste_per_cap < 3.5)) + 
         geom_point(aes(x = coastpror, y = plastic_waste_per_cap, color = continent)) + 
  stat_smooth(aes(x = coastpror, y = plastic_waste_per_cap)) + 
         labs(title = "Plastic waste vs. coastal population proportion" , 
                subtitle = "by continent" , 
                x = "Coastal population proportion (Coastal / total population)" , 
                y = "Plastic waste per capita" ) 

ggplot(data = plastic_waste %>%
         filter(plastic_waste_per_cap < 3.5), 
       mapping = aes(x = coastpror, y = plastic_waste_per_cap, color = continent)) +
  geom_point() + stat_smooth() +
         labs(title = "Plastic waste vs. coastal population proportion" , 
                subtitle = "by continent" , 
                x = "Coastal population proportion (Coastal / total population)" , 
                y = "Plastic waste per capita" ) 
```


